---
title: "ODE_MODEL"
author: "R.J. Beck"
date: "28/06/2020"
output: html_document
---

To perform the parameter estimation (which takes long) run the script fit_selected_v3b.R, which will save the output in the directory called fitted_output

```{r setup, include=FALSE}
base_dir <- "/home/richard/Documents/03_CTLsGoesNative/000_archive"
require(ggplot2)
require(gridExtra)
require(gdata)
require(deSolve)
knitr::opts_knit$set(root.dir = "/home/richard/Documents/03_CTLsGoesNative/000_archive/rscripts/parameter_estimation/v3")
source("process_fits_v3.R")
source("sensitivity_v3.R")
source("fit_model_v3b.R") ## contains the get.g and get.iv functions
ff <- list.files("fitted_output")  
if(FALSE){
pops <- lapply(ff,evaluate_members,fit_dir="fitted_output")
saveRDS(pops,"pops.rds")
}
x3 <- lapply(ff,evaluate_fits,fit_dir="fitted_output",d7s=FALSE)
x7 <- lapply(ff,evaluate_fits,fit_dir="fitted_output",d7s=TRUE)



s3_ke <- lapply(ff,vary_par ,fit_dir="fitted_output",d7s=FALSE)
s7_ke <- lapply(ff,vary_par ,fit_dir="fitted_output",d7s=TRUE)

s3_kq <- lapply(ff,vary_par ,fit_dir="fitted_output",vary="kq",d7s=FALSE)
s7_kq <- lapply(ff,vary_par ,fit_dir="fitted_output",vary="kq",d7s=TRUE)

```

Sensitivity analysis: 

1) Find best fitting parameters
2) Quantify the gradient of the solution w.r.t. each variable input parameter. 

Conclusion is that the model is identifiable (locally, anyway). This means that we don't need to consider other values for the parameters so much, because they would not lead to such good fits to the experimental data. Note that dq is likely so flat for the control case because not enough tumour cells became quiescent, for their recovery rate to matter so much.

```{r}
require(ggplot2)
fit_dir <- "fitted_output"
f5 <- ff[grep("g-5",ff)]

x <- sensitivity.analysis(ff=f5[1],fit_dir,range=seq(0.7,1.5,0.05))
y <- sensitivity.analysis(ff=f5[6],fit_dir,range=seq(0.7,1.5,0.05))

z <- rbind(x,y)

z$pars_f <- factor(z$varied,levels=c("s1","ki","di","kq","dq","ke","kr"))
z$pars_fe <- factor(z$pars_f,labels=c(expression(s),expression(k[i]),expression(d[i]),expression(k[q]),expression(d[q]),expression(k[e]),expression(k[r])))

z$treatment[z$treatment=="mAB"] <- "ACT+mAb"
z$treatment[z$treatment=="control"] <- "ACT-only"


px <- ggplot(z,aes(x=range,y=errors))+
  facet_grid(cols=vars(pars_fe),rows=vars(treatment),scales = "free",labeller=label_parsed)+
  geom_line()+
  scale_x_continuous("parameter value (relative to optimal)")+
  scale_y_continuous("RMSE")+
    theme_classic(base_size = 10)
px

ggsave("/home/richard/Documents/03_CTLsGoesNative/000_archive/figures/s5.tiff",plot = px,width = unit(7.5,"in"),height=unit(5,"in"),dpi=300)

```

```{r}

x <- readRDS(paste("stored_populations", list.files("stored_populations")[1],sep="/"))

fn.control <- x$fn.control
pops <- x$member$storepop

selected <- (1:10)*10-1

pops <- mclapply(selected, function(i){
  pop <- pops[[i]]
  errors <- evaluate_population(pop,fn.control)
  pop <- data.frame(pop)
  pop$error = errors
  pop$generation=i
  return(pop)
  
}, mc.cores=10)

pops <- do.call(rbind,pops)

pops <- pops[order(pops$error),]
pops <- head(pops,nrow(pops)/50)
```

```{r}
require(umap)
pops.df <- do.call(rbind,lapply(pops,process.population))
pops.df <- pops.df[pops.df$g==0.5,]

pop.data <- pops.df[,1:7]

pop.umap <- umap(pop.data)

umap.data <- data.frame(pop.umap$layout)

umap.data <- cbind(umap.data,pops.df[,8:ncol(pops.df)])

p <- ggplot(umap.data,aes(x=X1,y=X2,color=treatment))+
  geom_point()+
  theme_classic()
p

p <- ggplot(umap.data,aes(x=X1,y=X2,color=id))+
  geom_point()+
  theme_classic()
p

p <- ggplot(subset(umap.data,treatment=="control"),aes(x=X1,y=X2,color=errors))+
  geom_point()+
  theme_classic()
p

p <- ggplot(subset(umap.data,treatment=="mAB"),aes(x=X1,y=X2,color=errors))+
  geom_point()+
  theme_classic()
p


```

```{r}

treatments <- sapply(pops,function(p) p$treatment)
mab <- pops[treatments=="mAB"]
con <- pops[treatments=="control"]

mab.err <- unlist(lapply(mab,function(x) x$error))
hist(mab.err)
```

INVESTIGATE THE BEST FITTING PARAMETERS

```{r, echo=FALSE, message=FALSE,warn=FALSE}
pardf <- data.frame(do.call(rbind,lapply(x3, function(h) h$par)))
pardf <- reshape2::melt(pardf,id.vars=c("g","const_inf","dr","id","treatment","QSS","t_in","T0","t50"))
pardf <- pardf[pardf$variable!="pow",]
pardf$value <- as.numeric(pardf$value)
pardf <- pardf[pardf$value>0,]

par_summary <- aggregate(pardf[c("value")],by=pardf[c("g","variable","QSS","treatment")],mean)
par_summary$variable <- factor(par_summary$variable,levels=c("s1","ki","di","kq","dq","ke","kr"))
pardf$variable <- factor(pardf$variable,levels=c("s1","ki","di","kq","dq","ke","kr"))




pa <- ggplot(subset(pardf,g%in%c(0.4,0.5,0.6)),aes(x=variable,y=value,color=treatment,shape=treatment))+
  facet_grid(cols=vars(g))+
  geom_jitter(height=0,width=0.3)+
  geom_point(data=subset(par_summary,g%in%c(0.4,0.5,0.6)),shape=95,size=10)+
  scale_x_discrete("parameter",labels=c(expression(s),expression(k[i]), expression(d[i]), expression(k[q]),
                                        expression(d[q]), expression(k[e]),expression(k[r])))+
  scale_y_log10(expression(fitted~rate~(day^-1)))+
  theme_classic(base_size = 12)+
    scale_color_discrete(labels=c("ACT-only","ACT+mAb"))+
  scale_shape_discrete(labels=c("ACT-only","ACT+mAb"))+
  theme(legend.position = "none")
pa


pa0 <- ggplot(subset(pardf,g==0.5),aes(x=variable,y=value,color=treatment,shape=treatment))+
  #facet_grid(cols=vars(g))+
  geom_jitter(height=0,width=0.3)+
  geom_point(data=subset(par_summary,g==0.5),shape=95,size=8)+
  scale_x_discrete("parameter",labels=c(expression(s),expression(k[i]), expression(d[i]), expression(k[q]),
                                        expression(d[q]), expression(k[e]),expression(k[r])))+
  scale_y_log10(expression(fitted~rate~(day^-1)))+
  theme_classic(base_size = 10)+
  scale_color_discrete(labels=c("ACT-only","ACT+mAb"))+
  scale_shape_discrete(labels=c("ACT-only","ACT+mAb"))
pa0

f3c <- egg::ggarrange(pa0,nrow=1,labels=c("C"))
ggsave("/home/richard/Documents/03_CTLsGoesNative/000_archive/figures/f3c.tiff",plot = f3c,width = unit(3.5,"in"),height=unit(2.25,"in"),dpi=300)

print(subset(par_summary,g==0.5))

```
lowest error is for parms with id=012 (control) and id = 043 (mAB). This info can be used to recover the best fitting parms

```{r}
bestfits <- x3[sapply(x3, function(h) h$errordf$id%in%c("012","043"))]

bestfits[[1]]$par
bestfits[[2]]$par

```


```{r}
errordf <- data.frame(do.call(rbind,lapply(x3, function(h) h$errordf)))
pe0 <- ggplot(subset(errordf,g%in%c(0.4,0.5,0.6)),aes(x=treatment,y=error,color=treatment,shape=treatment))+
  facet_grid(cols=vars(g))+
  geom_jitter(height=0,width=0.3)+
  scale_x_discrete("treatment")+
  scale_y_continuous("RMSE")+
  theme_classic(base_size = 12)+
  theme(legend.position = "none")
pe0

#pe1 <- ggplot(subset(errordf,QSS==1),aes(x=g,y=error,color=treatment,shape=treatment))+
 # facet_grid(rows=vars(QSS),labeller = label_bquote(rows=QSS==.(QSS)))+
  #geom_jitter(height=0,width=0.3)+
  #scale_x_discrete("growth rate (g)")+
  #scale_y_continuous("RMSE")+
  #theme_classic(base_size = 12)
#pe1

```

```{r}
ivdf <- data.frame(do.call(rbind,lapply(x3, function(h) h$df)))
ivdf <- reshape2::melt(ivdf,id.vars=c("treatment","id","time","QSS"))

idf <- rbind(get.iv("mAB"),get.iv("control"))
idf <- reshape2::melt(idf,id.vars=c("day","ID"))
names(idf)[names(idf)=="ID"] <- "treatment"

pb0 <- ggplot(subset(ivdf,variable%in%c("TCm","CTLm","CTLa","killing")&QSS==0),aes(x=as.numeric(time),y=as.numeric(value),color=treatment))+
  facet_grid(rows=vars(variable),scales = "free_y")+
#  geom_segment(data=mdf,aes(x=t1,xend=t2,y=value,yend=value))+
  geom_line(aes(group=id),size=1.2)+
  geom_jitter(data=subset(idf,variable%in%c("TCm","CTLm","CTLa","killing")), aes(x=as.numeric(day),y=as.numeric(value),shape=treatment),height=0,width=0.3)+
  scale_x_continuous("day", breaks=seq(0,15,3))+
  scale_y_continuous(expression(paste("process rate", (day^-1))))+
  theme_classic(base_size = 12)
pb0

pc0 <- ggplot(subset(ivdf,variable%in%c("ET")&QSS==0),aes(x=as.numeric(time),y=as.numeric(value),color=treatment))+
  geom_line(aes(group=id),size=1.2)+
  geom_jitter(data=subset(idf,variable%in%c("ET")), aes(x=as.numeric(day),y=as.numeric(value),shape=treatment),height=0,width=0.3)+
  scale_x_continuous("day", breaks=seq(0,15,3))+
  scale_y_continuous(expression(paste("ET ratio", (day^-1))))+
  theme_classic(base_size = 12)
pc0


```

Both fits to volumetric growth data (d3s and d7s)

```{r}

require(ggplot2)
library(gridExtra)
library(cowplot)
get_legend<-function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
}

day_labels <- c('d3.s'="day 3", 'd7.s'="day 7")

ivdf3 <- data.frame(do.call(rbind,lapply(x3, function(h) h$df)))
ivdf3$start <- "d3.s"
ivdf3$net_g <- ivdf3$TCm-ivdf3$killing*ivdf3$ET

ivdf7 <- data.frame(do.call(rbind,lapply(x7, function(h) h$df)))
ivdf7$start <- "d7.s"
ivdf7$net_g <- ivdf7$TCm-ivdf7$killing*ivdf7$ET

ivdf <- rbind(ivdf3,ivdf7)

ivdf <- reshape2::melt(ivdf,id.vars=c("treatment","id","time","start","g"))

mdf <- rbind(get.g("mAB"),get.g("control"))
names(mdf)[names(mdf)=="g"] <- "value"
mdf$variable <- "net_g"
mdf$day <- mdf$t1+(mdf$t2-mdf$t1)/2
mdf_summ <- aggregate(mdf[c("value")],by=mdf[c("treatment","variable","day","start")],mean)
mdf_summ_sd <- aggregate(mdf[c("value")],by=mdf[c("treatment","variable","day","start")],sd)
mdf_summ$sd <- mdf_summ_sd$value


idf_d7 <- get.d7.data("mAB")
idf_d7$treatment <- "mAB"
tmp <- get.d7.data("control")
tmp$treatment <-"control"
idf_d7 <- rbind(idf_d7,tmp)
idf_d7$day <- 10
idf_d7 <- reshape2::melt(idf_d7,id.vars=c("day","treatment"))
idf_d7$start <- "d7.s"

idf <- rbind(get.iv("mAB"),get.iv("control"))
idf <- reshape2::melt(idf,id.vars=c("day","ID"))
names(idf)[names(idf)=="ID"] <- "treatment"
idf$start <- "d3.s"

idf <- rbind(idf,idf_d7)

idf_summ <- aggregate(idf[c("value")],by=idf[c("treatment","variable","day","start")],mean)
idf_summ_sd <- aggregate(idf[c("value")],by=idf[c("treatment","variable","day","start")],sd)
idf_summ$sd <- idf_summ_sd$value

ivd3 <- ggplot(subset(ivdf,variable%in%c("TCm","CTLm","CTLa","killing")&g==0.5&start=="d3.s"),aes(x=as.numeric(time),y=as.numeric(value),color=treatment))+
  facet_grid(rows=vars(variable),scales = "free_y")+
#  geom_segment(data=mdf,aes(x=t1,xend=t2,y=value,yend=value))+
  geom_line(aes(group=id),size=1.2)+
  geom_point(data=subset(idf_summ,variable%in%c("TCm","CTLm","CTLa","killing")&start=="d3.s"), aes(x=as.numeric(day),y=value))+
  geom_errorbar(data=subset(idf_summ,variable%in%c("TCm","CTLm","CTLa","killing")&start=="d3.s"), aes(x=as.numeric(day),ymin=value-sd,ymax=value+sd))+
  
  scale_x_continuous("day", breaks=seq(0,15,3))+
  scale_y_continuous(expression(paste("process rate", (day^-1))))+
  theme_classic(base_size = 12)+
  scale_color_discrete(labels=c("ACT-only","ACT+mAb"))+
  guides(color=FALSE)
ivd3

pb3_all <- ggplot(subset(ivdf,variable%in%c("TCm","CTLm","CTLa","killing")&start=="d3.s"&g%in%c(0.4,0.5,0.6)),aes(x=as.numeric(time),y=as.numeric(value),color=treatment))+
  facet_grid(rows=vars(variable),cols=vars(g),scales = "free_y")+
  geom_line(aes(group=id),size=1.2)+
  geom_point(data=subset(idf_summ,variable%in%c("TCm","CTLm","CTLa","killing")&start=="d3.s"), aes(x=as.numeric(day),y=value))+
  geom_errorbar(data=subset(idf_summ,variable%in%c("TCm","CTLm","CTLa","killing")&start=="d3.s"), aes(x=as.numeric(day),ymin=value-sd,ymax=value+sd))+
  scale_x_continuous("day", breaks=seq(0,15,3))+
  scale_y_continuous(expression(paste("process rate", (day^-1))))+
  theme_classic(base_size = 12)+
  scale_color_discrete(labels=c("ACT-only","ACT+mAb"))+
  theme(legend.position = "top")
pb3_all

pb7_all <- ggplot(subset(ivdf,variable%in%c("TCm","CTLm","CTLa","killing")&start=="d7.s"&g%in%c(0.4,0.5,0.6)),aes(x=as.numeric(time),y=as.numeric(value),color=treatment))+
  facet_grid(rows=vars(variable),cols=vars(g),scales = "free_y")+
  geom_line(aes(group=id),size=1.2)+
  geom_point(data=subset(idf_summ,variable%in%c("TCm","CTLm","CTLa","killing")&start=="d7.s"), aes(x=as.numeric(day),y=value))+
  geom_errorbar(data=subset(idf_summ,variable%in%c("TCm","CTLm","CTLa","killing")&start=="d7.s"), aes(x=as.numeric(day),ymin=value-sd,ymax=value+sd))+
  
  scale_x_continuous("day", breaks=seq(0,15,3))+
  scale_y_continuous(expression(paste("process rate", (day^-1))))+
  theme_classic(base_size = 12)+
  scale_color_discrete(labels=c("ACT-only","ACT+mAb"))+
  theme(legend.position = "none")
pb7_all

ivd7 <- ggplot(subset(ivdf,variable%in%c("TCm","CTLm","CTLa","killing")&g==0.5&start=="d7.s"),aes(x=as.numeric(time),y=as.numeric(value),color=treatment))+
  facet_grid(rows=vars(variable),scales = "free_y")+
  geom_point(data=subset(idf_summ,variable%in%c("TCm","CTLm","CTLa","killing")&start=="d7.s"), aes(x=as.numeric(day),y=value))+
  geom_errorbar(data=subset(idf_summ,variable%in%c("TCm","CTLm","CTLa","killing")&start=="d7.s"), aes(x=as.numeric(day),ymin=value-sd,ymax=value+sd))+
  geom_line(aes(group=id),size=1.2)+
  scale_x_continuous("day", breaks=seq(0,15,3))+
  scale_y_continuous(expression(paste("process rate", (day^-1))))+
  theme_classic(base_size = 12)+
  scale_color_discrete(labels=c("ACT-only","ACT+mAb"))+
  guides(color=FALSE)
ivd7

etd3 <- ggplot(subset(ivdf,variable%in%c("ET")&g==0.5&start=="d3.s"),aes(x=as.numeric(time),y=as.numeric(value),color=treatment))+
  #facet_grid(cols=vars(start),scales = "free_y")+
  geom_line(aes(group=id),size=1.2)+
  geom_point(data=subset(idf_summ,variable%in%c("ET")&start=="d3.s"), aes(x=as.numeric(day),y=value))+
  geom_errorbar(data=subset(idf_summ,variable%in%c("ET")&start=="d3.s"), aes(x=as.numeric(day),ymin=value-sd,ymax=value+sd))+
  scale_x_continuous("day", breaks=seq(0,15,3))+
  scale_y_continuous(expression(paste("ET ratio")))+
  theme_classic(base_size = 12)+
  scale_color_discrete(labels=c("ACT-only","ACT+mAb"))+
  guides(color=FALSE)
etd3

pc_all <- ggplot(subset(ivdf,variable%in%c("ET")&g%in%c(0.4,0.5,0.6)),aes(x=as.numeric(time),y=as.numeric(value),color=treatment))+
  facet_grid(cols=vars(g),rows=vars(start),scales = "free_y",labeller = labeller(start=day_labels))+
  geom_line(aes(group=id),size=1.2)+
  geom_point(data=subset(idf_summ,variable%in%c("ET")), aes(x=as.numeric(day),y=value))+
  geom_errorbar(data=subset(idf_summ,variable%in%c("ET")), aes(x=as.numeric(day),ymin=value-sd,ymax=value+sd))+
  scale_x_continuous("day", breaks=seq(0,15,3))+
  scale_y_continuous(expression(paste("ET ratio")))+
  theme_classic(base_size = 12)+
  scale_color_discrete(labels=c("ACT-only","ACT+mAb"))+
  theme(legend.position = "none")
pc_all

etd7 <- ggplot(subset(ivdf,variable%in%c("ET")&g==0.5&start=="d7.s"),aes(x=as.numeric(time),y=as.numeric(value),color=treatment))+
  geom_line(aes(group=id),size=1.2)+
  geom_point(data=subset(idf_summ,variable%in%c("ET")&start=="d7.s"), aes(x=as.numeric(day),y=value))+
  geom_errorbar(data=subset(idf_summ,variable%in%c("ET")&start=="d7.s"), aes(x=as.numeric(day),ymin=value-sd,ymax=value+sd))+
  scale_x_continuous("day", breaks=seq(0,15,3))+
  scale_y_continuous(expression(paste("ET ratio")))+
  theme_classic(base_size = 12)+
  scale_color_discrete(labels=c("ACT-only","ACT+mAb"))+
  guides(color=FALSE)
etd7



gd3 <- ggplot(subset(ivdf,variable%in%c("net_g")&g==0.5&start=="d3.s"),aes(x=as.numeric(time),y=as.numeric(value),color=treatment))+
  #facet_grid(cols=vars(start),scales = "free_y")+
  geom_point(data=subset(mdf_summ,start=="d3.s"), aes(x=as.numeric(day),y=value))+
  geom_errorbar(data=subset(mdf_summ,start=="d3.s"), aes(x=as.numeric(day),ymin=value-sd,ymax=value+sd))+
  geom_line(aes(group=id),size=1.2)+
  scale_x_continuous("day", breaks=seq(0,15,3))+
  scale_y_continuous(expression(paste("net growth", (day^-1))))+
  theme_classic(base_size = 12)+
  scale_color_discrete(labels=c("ACT-only","ACT+mAb"))+
  guides(color=FALSE)
gd3

gd7 <- ggplot(subset(ivdf,variable%in%c("net_g")&g==0.5&start=="d7.s"),aes(x=as.numeric(time),y=as.numeric(value),color=treatment))+
  #facet_grid(cols=vars(start),scales = "free_y")+
  geom_point(data=subset(mdf_summ,start=="d7.s"), aes(x=as.numeric(day),y=value))+
  geom_errorbar(data=subset(mdf_summ,start=="d7.s"), aes(x=as.numeric(day),ymin=value-sd,ymax=value+sd))+
  geom_line(aes(group=id),size=1.2)+
  scale_x_continuous("day", breaks=seq(0,15,3))+
  scale_y_continuous(expression(paste("net growth", (day^-1))))+
  theme_classic(base_size = 12)+
  scale_color_discrete(labels=c("ACT-only","ACT+mAb"))+
  guides(color=FALSE)
gd7

pd_all <- ggplot(subset(ivdf,variable%in%c("net_g")&g%in%c(0.4,0.5,0.6)),aes(x=as.numeric(time),y=as.numeric(value),color=treatment))+
  facet_grid(rows=vars(start),cols=vars(g),scales = "free_y",labeller = labeller(start=day_labels))+
  geom_point(data=mdf_summ, aes(x=as.numeric(day),y=value))+
  geom_errorbar(data=mdf_summ, aes(x=as.numeric(day),ymin=value-sd,ymax=value+sd))+
  geom_line(aes(group=id),size=1.2)+
  scale_x_continuous("day", breaks=seq(0,15,3))+
  scale_y_continuous(expression(paste("net growth", (day^-1))))+
  theme_classic(base_size = 12)+
  scale_color_discrete(labels=c("ACT-only","ACT+mAb"))+
  theme(legend.position = "bottom",legend.direction = "horizontal")

p5_legend <- get_legend(pd_all)
pd_all <- pd_all +theme(legend.position = "none")

pd_all



legend_height <- 0.05
label_height<-0.03
r1 <- 0.45
r2 <- 0.225
r3 <- r2

f5 <- ggdraw()+
   draw_plot(ivd3, 0, legend_height+r2+r3, .5, r1)+ 
 draw_plot(ivd7, 0.5, legend_height+r2+r3, .5, r1)+ 
  draw_plot(etd3, 0, legend_height+r3, .5, r2) +
   draw_plot(etd7, 0.5, legend_height+r3, .5, r2)+ 
    draw_plot(gd3, 0, legend_height, .5, r3)+ 
     draw_plot(gd7, 0.5, legend_height, .5, r3)+ 
      draw_plot(p5_legend, 0, 0, 1, legend_height) +
  draw_plot_label(c("A","B","C","D","E","F"), c(0, 0.5, 0, 0.5, 0, 0.5), c(1-label_height,1-label_height,legend_height+r2+r3,legend_height+r2+r3,legend_height+r3,legend_height+r3), size = 15)+
    draw_plot_label(c("day 3 start", "day 7 start"), x=c(0.25,0.75),y=c(1,1),hjust = c(0.5,0.5))
ggsave("/home/richard/Documents/03_CTLsGoesNative/000_archive/figures/f5.tiff",plot = f5,width = unit(7.5,"in"),height=unit(10,"in"),dpi=300)

s3_fig_left <- egg::ggarrange(pd_all,pc_all,pa,pe0,nrow=4,labels=c("A","B","C","D"))
s3_fig_right <- egg::ggarrange(pb3_all,pb7_all,nrow=2,labels=c("E","F"))

s3_fig <- grid.arrange(s3_fig_left,s3_fig_right,nrow=1)
ggsave("/home/richard/Documents/03_CTLsGoesNative/000_archive/figures/s3.tiff",plot = s3_fig,width = unit(10,"in"),height=unit(7.5,"in"),dpi=300)
```
Predict the infiltration into each window.

```{r}
tumdf3 <- data.frame(do.call(rbind,lapply(x3, function(h) h$out)))
tumdf3$start <- "d3"
tumdf3$inf <- (tumdf3$Tc+tumdf3$Tq)^(2/3)*tumdf3$s1/(tumdf3$Tc+tumdf3$Tq)
tumdf3$inf[tumdf3$time<3]<-0
tumdf7 <- data.frame(do.call(rbind,lapply(x7, function(h) h$out)))
tumdf7$start <- "d7"
tumdf7$inf <- (tumdf7$Tc+tumdf7$Tq)^(2/3)*tumdf7$s1/(tumdf7$Tc+tumdf7$Tq)
tumdf7$inf[tumdf7$time<7]<-0
tumdf <- rbind(tumdf3,tumdf7)

pi3 <- ggplot(tumdf,aes(x=time,y=10^3*inf/24,color=treatment,group=id))+
  facet_grid(rows=vars(start),cols=vars(g),scales = "free_y")+
  geom_line(aes(group=id))+
  scale_x_continuous("day", breaks=seq(0,15,3))+
  scale_y_continuous(expression(paste("infiltration", (hour^-1%.%TCx10^-3))))+
  theme_classic(base_size = 20)
pi3

```

fit to volume data

```{r}

vdf <- read.csv(paste(base_dir,"/data/formattedGrowthAll.csv",sep=""))
idf <- read.csv(paste(base_dir,"/data/formattedIntravital.csv",sep=""))

tc.density <- mean(idf$TC)/(0.1*0.35^2) #126393.9

vdf <- vdf[vdf$treatment%in%c("mAB","control"),]
vdf$start[vdf$start=="d3.s"] <- "d3"
vdf$start[vdf$start=="d7.s"] <- "d7"

tumdf3 <- data.frame(do.call(rbind,lapply(x3, function(h) h$out)))
tumdf3$start <- "d3"
tumdf3$inf <- (tumdf3$Tc+tumdf3$Tq)^(2/3)*tumdf3$s1/(tumdf3$Tc+tumdf3$Tq)
tumdf3$inf[tumdf3$time<3]<-0
tumdf7 <- data.frame(do.call(rbind,lapply(x7, function(h) h$out)))
tumdf7$start <- "d7"
tumdf7$inf <- (tumdf7$Tc+tumdf7$Tq)^(2/3)*tumdf7$s1/(tumdf7$Tc+tumdf7$Tq)
tumdf7$inf[tumdf7$time<7]<-0
tumdf <- rbind(tumdf3,tumdf7)

tumdf <- subset(tumdf,g==0.5)

tumdf$V <- (tumdf$Tc+tumdf$Tq)/tc.density

treatment_labels <- c(control="ACT-only",mAB="ACT+mAB")
day_labels <- c(d3="day 3", d7="day 7")

pV <- ggplot(tumdf,aes(x=time,y=V))+
  facet_grid(rows=vars(start),cols=vars(treatment),scales = "free_y",labeller = labeller(start=day_labels,treatment=treatment_labels))+
  geom_line(aes(group=id))+
  geom_point(data=vdf,aes(x=day,y=vol))+
  scale_x_continuous("day", breaks=seq(0,15,3))+
  scale_y_log10(expression(paste("tumour volume", (mm^3))))+
  theme_classic(base_size = 14)
pV

ggsave("/home/richard/Documents/03_CTLsGoesNative/000_archive/figures/s4.tiff",plot = pV,width = unit(7.5,"in"),height=unit(5,"in"),dpi=300)

```

```{r}

# process tumour data by removing rows for tumours that get smaller than initial value (in violation of our tumour cell saturation assumption)
tumour.progress <- function(x){
  x <- subset(x,g==0.5)
  x$Ttot <- x$Tc+x$Tq
  x$Ttot[x$Ttot<1200]<- NA
  return(x)
}

x <- tumour.progress(do.call(rbind,s7_ke))
y <- tumour.progress(do.call(rbind,s7_kq))





# 126393.9 is tc.density 
p4ke <- ggplot(x,aes(x=time,y=(Ttot)/126393.9,group=id,color=treatment))+
  facet_grid(cols=vars(ratio),labeller=label_bquote(cols=k[e]==.(ratio)~widehat(k[e]) ))+
  geom_line()+
  scale_x_continuous("day", breaks=seq(0,15,3),limits=c(0,15))+
  scale_y_log10(expression(paste("tumour volume", (mm^3))),limits=c(0.008,3))+
  theme_classic(base_size = 20)+
  theme(text = element_text(size=12))+
  scale_color_discrete(labels=c("ACT-only","ACT+mAb"))+
  guides(color=FALSE)
p4ke

p <- ggplot(x,aes(x=time,y=E/(Tc+Tq),group=id,color=treatment))+
  facet_grid(cols=vars(ratio))+
  geom_line()+
  scale_y_continuous(limits=c(0,.5))
p


p4kq <- ggplot(y,aes(x=time,y=Ttot/126393.9,group=id,color=treatment))+
  facet_grid(cols=vars(ratio),labeller=label_bquote(cols=k[q]==.(ratio)~widehat(k[q]) ))+
  geom_line()+
  scale_x_continuous("day", breaks=seq(0,15,3),limits=c(0,15))+
  scale_y_log10(expression(paste("tumour volume", (mm^3))),limits=c(0.008,15))+
  theme_classic(base_size = 20)+
  scale_color_discrete(labels=c("ACT-only","ACT+mAb"))+
  theme(legend.position = "bottom",text=element_text(size=12))
p4kq


p <- ggplot(y,aes(x=time,y=E/(Tc+Tq),group=id,color=treatment))+
  facet_grid(cols=vars(ratio))+
  geom_line()+
  scale_y_continuous()
p

f4 <- egg::ggarrange(p4ke,p4kq,nrow=2,labels=c("A","B"))
ggsave("/home/richard/Documents/03_CTLsGoesNative/000_archive/figures/f4.tiff",plot = f4,width = unit(7.5,"in"),height=unit(6.5,"in"),dpi=300)
```
