---
title: "Analysis of Experimental Data"
author: "R.J. Beck"
date: "5/18/2020"
output: html_document
---

This script generates all figures that are derived directly from the experimental data i.e.:

Figures 1-2
Figures S1-S2

```{r,setup}

base_dir <- "/home/richard/Documents/03_CTLsGoesNative/000_archive"
data_proc_funcs <- paste(base_dir,"/rscripts/utility/data_processing_functions.R",sep="")
source (data_proc_funcs)

require(ggplot2)
mouse_shapes <- c(0,1,2,5,12,13,4,9)
names(mouse_shapes) <- c(49,64,65,57,50,66,67,55)
mouse_colors <- c(1,2,3,4,1,2,3,4)
names(mouse_colors) <- c(49,64,65,57,50,66,67,55)
iv.proc <- get.iv()

dayz <- c("green", "blue")

dfstat <- aggregate(iv.proc[c("TCa","TCm")],by=iv.proc[c("ID","day","mouse")],mean)
dfstat1 <- aggregate(iv.proc[c("CTLa","CTLm")],by=iv.proc[c("ID","day","mouse")],mean)
dfstat$type <- "tumour"
dfstat1$type <- "CTL"
names(dfstat)[names(dfstat)%in%c("TCa","TCm")] <- c("apt","mit")
names(dfstat1)[names(dfstat1)%in%c("CTLa","CTLm")] <- c("apt","mit")
dfstat <- rbind(dfstat,dfstat1)
dfstat$net <- dfstat$mit-dfstat$apt

dfvec <- dfstat[dfstat$mouse!="m1",]
dfvec$uid <- interaction(dfvec$ID,dfvec$mouse,dfvec$type)


for(id in unique(dfvec$uid)){ ## get difference between day 6 and 9
  dfvec$apt[dfvec$uid==id] <- dfvec$apt[dfvec$uid==id]-dfvec$apt[dfvec$uid==id&dfvec$day==6]
  dfvec$mit[dfvec$uid==id] <- dfvec$mit[dfvec$uid==id]-dfvec$mit[dfvec$uid==id&dfvec$day==6]
}

ivstat <- iv.proc[,c("ID","mouse","day","TCa","TCm","TC","CTL")]
ivstat1 <- iv.proc[,c("ID","mouse","day","CTLa","CTLm","TC","CTL")]
ivstat$type <- "tumour"
ivstat1$type <- "CTL"
names(ivstat)[names(ivstat)%in%c("TCa","TCm")] <- c("apt","mit")
names(ivstat1)[names(ivstat1)%in%c("CTLa","CTLm")] <- c("apt","mit")
ivstat <- rbind(ivstat,ivstat1)
ivstat$net <- ivstat$mit-ivstat$apt
ivstat$cell_count <- ivstat$TC
ivstat$cell_count[ivstat$type=="CTL"]<-ivstat$CTL[ivstat$type=="CTL"]



x <- lapply(1:nrow(ivstat),function(i){
  x <- ivstat[i,]
  y <-subset(dfstat,mouse==x$mouse&type==x$type&day==x$day)
  return(data.frame(APT=y$apt,MIT=y$mit))
  })
x <- do.call(rbind,x)
ivstat<- cbind(ivstat,x)
ivstat$ID[ivstat$ID=="control"] <- "ACT-only"
ivstat$ID[ivstat$ID=="mAB"] <- "ACT+mAb"
dfstat$ID[dfstat$ID=="control"] <- "ACT-only"
dfstat$ID[dfstat$ID=="mAB"] <- "ACT+mAb"
dfvec$ID[dfvec$ID=="control"] <- "ACT-only"
dfvec$ID[dfvec$ID=="mAB"] <- "ACT+mAb"
```


```{r}
require(ggplot2)
library(gridExtra)
library(cowplot)
get_legend<-function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
}

p1 <- ggplot(subset(ivstat,type=="tumour"), aes(x=apt, y=mit))+
  facet_grid(cols=vars(ID))+
  geom_point(aes(shape=as.character(mouse),color=as.character(day)),size=2)+
  geom_abline(slope=1,intercept=0,color="red",linetype="dashed")+
  geom_point(data=subset(dfstat,type=="tumour"),aes(shape=as.character(mouse),color=as.character(day)),size=4,show.legend = FALSE)+
  geom_line(data=subset(dfstat,type=="tumour"),aes(group=as.character(mouse)),show.legend = FALSE)+
  scale_x_continuous(expression(paste("TC apoptosis ",(TC^-1~day^-1))))+
  scale_y_continuous(expression(paste("TC mitosis ",(TC^-1~day^-1))))+
  scale_color_manual("day",values=dayz)+
  scale_shape_manual("mouse",values=mouse_shapes)+
  theme_classic(base_size = 10)+
  theme(legend.position = "top",legend.direction = "horizontal")
  ab_legend <- get_legend(p1)
p1 <- p1  +  theme(legend.position = "none")


p1a <- ggplot(subset(ivstat,type=="tumour"), aes(x=apt, y=mit))+
  facet_grid(cols=vars(ID))+
  geom_segment(aes(xend=APT,yend=MIT,color=as.character(mouse)))+
  geom_point(aes(color=as.character(mouse),size=cell_count))+
  scale_x_continuous(expression(paste("TC apoptosis ",(TC^-1~day^-1))))+
  scale_y_continuous(expression(paste("TC mitosis ",(TC^-1~day^-1))))+
  scale_color_viridis_d("mouse")+
  scale_size_area("TCs",limits=c(0,2500))+
  theme_classic(base_size = 10)+
  theme(legend.position = "none")



p2 <- ggplot(subset(ivstat,type=="CTL"), aes(x=apt, y=mit))+
 facet_grid(cols=vars(ID))+
  geom_point(aes(shape=as.character(mouse),color=as.character(day)),size=2)+
  geom_abline(slope=1,intercept=0,color="red",linetype="dashed")+
  geom_point(data=subset(dfstat,type=="CTL"),aes(shape=as.character(mouse),color=as.character(day)),size=4,show.legend = FALSE)+
  geom_line(data=subset(dfstat,type=="CTL"),aes(group=as.character(mouse)))+
  scale_x_continuous(expression(paste("CTL apoptosis ",(CTL^-1~day^-1))))+
  scale_y_continuous(expression(paste("CTL mitosis ",(CTL^-1~day^-1))))+
  scale_color_manual("day",values=dayz)+
  scale_shape_manual("mouse",values=mouse_shapes)+
  scale_linetype_discrete("")+
  theme_classic(base_size = 10)+
  theme(legend.position = "none")

p2a <- ggplot(subset(ivstat,type=="CTL"), aes(x=apt, y=mit))+
 facet_grid(cols=vars(ID))+
  geom_segment(aes(xend=APT,yend=MIT,color=as.character(mouse)))+
  geom_point(aes(color=as.character(mouse),size=cell_count))+
  scale_x_continuous(expression(paste("CTL apoptosis ",(CTL^-1~day^-1))))+
  scale_y_continuous(expression(paste("CTL mitosis ",(CTL^-1~day^-1))))+
  scale_color_viridis_d("mouse")+
  scale_size_area("cells",limits=c(0,2000))+
  theme_classic(base_size = 10)+
  theme(legend.position = "top",legend.direction = "horizontal")
  cd_legend <- get_legend(p2a)
p2a <- p2a +theme(legend.position="none")
  
p3 <- ggplot(subset(dfvec,type=="CTL"), aes(x=apt, y=mit))+
  facet_grid(cols=vars(ID))+
    geom_abline(slope=1,intercept=0,color="red",linetype="dashed")+
  geom_abline(slope=-1,intercept=0,color="red",linetype="dashed")+
  geom_point(aes(shape=as.character(mouse),color=as.character(day)),size=4)+
  geom_line(aes(group=as.character(mouse)))+
  scale_x_continuous(expression(paste("CTL apoptosis change",(d9-d6))))+
  scale_y_continuous(expression(paste("CTL mitosis change",(d9-d6))))+
  scale_color_manual("day",values=dayz)+
  scale_shape_manual("mouse",values=mouse_shapes)+
  scale_linetype_discrete("")+
  theme_classic(base_size = 10)+
  theme(legend.position = "top",legend.direction = "horizontal")
ef_legend <- get_legend(p3)
p3 <- p3 +theme(legend.position="none")

p4 <- ggplot(subset(dfvec,type=="tumour"), aes(x=apt, y=mit))+
  facet_grid(cols=vars(ID))+
  geom_abline(slope=1,intercept=0,color="red",linetype="dashed")+
  geom_point(aes(shape=as.character(mouse),color=as.character(day)),size=4)+
  geom_line(aes(group=as.character(mouse)))+
  scale_x_continuous(expression(paste("TC apoptosis change ",(d9-d6))))+
  scale_y_continuous(expression(paste("TC mitosis change ",(d9-d6))))+
  scale_color_manual("day",values=dayz)+
  scale_shape_manual("mouse",values=mouse_shapes)+
  scale_linetype_discrete("")+
  theme_classic(base_size = 10)+
  theme(legend.position = "none")
```

```{R}

legend_height<-0.15
label_height <- 0.07

pab <- ggdraw()+
draw_plot(ab_legend, 0, 0, 1, legend_height) +
draw_plot(p1, 0, legend_height, .5, 1-legend_height-label_height) +
draw_plot(p2, .5, legend_height, .5, 1-legend_height-label_height) +
draw_plot_label(c("A", "B", ""), c(0, 0.5, 0), c(1-label_height, 1-label_height, 1-legend_height), size = 15)+
  draw_plot_label(c("tumour cells", "CTLs"), x=c(0.25,0.75),y=c(1,1),hjust = c(0.5,0.5))
ggsave("/home/richard/Documents/03_CTLsGoesNative/000_archive/figures/f1ab.tiff",plot = pab,width = unit(7.5,"in"),height=unit(3.3,"in"),dpi=300)

pcd <- ggdraw() +
 draw_plot(cd_legend, 0, 0, 1, legend_height) +
 draw_plot(p1a, 0, legend_height, .5, 1-legend_height) +
  draw_plot(p2a, .5, legend_height, .5, 1-legend_height) +
  draw_plot_label(c("C", "D", ""), c(0, 0.5, 0), c(1, 1, 1-legend_height), size = 15)
ggsave("/home/richard/Documents/03_CTLsGoesNative/000_archive/figures/f1cd.tiff",plot = pcd,width = unit(7.5,"in"),height=unit(3,"in"),dpi=300)

pef <- ggdraw() +
 draw_plot(ef_legend, 0, 0, 1, legend_height) +
 draw_plot(p3, 0, legend_height, .5, 1-legend_height) +
  draw_plot(p4, .5, legend_height, .5, 1-legend_height) +
  draw_plot_label(c("E", "F", ""), c(0, 0.5, 0), c(1, 1, 1-legend_height), size = 15)
ggsave("/home/richard/Documents/03_CTLsGoesNative/000_archive/figures/f1ef.tiff",plot = pef,width = unit(7.5,"in"),height=unit(3,"in"),dpi=300)

```

```{R}
print("not rendering")
if(TRUE){
  p <- list(p1,p2,p3,p4)
  pad <- egg::ggarrange(p1,p2,p4,p3,nrow=2)
  pad_alliv <- egg::ggarrange(p1,p2,p1a,p2a,p4,p3,nrow=3,labels = c("A","B","C","D","E","F"))
  
}

ggsave("/home/richard/Documents/03_CTLsGoesNative/000_archive/figures/f1.tiff",plot = pad_alliv,width = unit(7.5,"in"),height=unit(10,"in"),dpi=300)

```


Tumour volume over time. First convert this to growth rate estimates, which comes with two benefits:

1) Second estimate for tumour growth. Can be compared to tumour mitosis rates to see if these are consistent. If not, suggests that the tumour is growing via "diffusive" means.
2) Avoids inaccuracies due to unknown volumetric density of tumour cells (results remain true so far as the density of the tumour remains constant, which the comparison with the mitosis data (1) reveals).

```{r, echo=FALSE}
require(ggplot2)

vol_meas_funcs <- paste(base_dir,"/rscripts/utility/measure_growth_rates.R",sep="")
source(vol_meas_funcs)


vdf <- read.csv(paste(base_dir,"/data/formattedGrowthAll.csv",sep=""))
mdf <- get_volgrowth(paste(base_dir,"/data/formattedGrowthAll.csv",sep=""))

relabl <- function(df){

  st <- df$start
  tr <- df$treatment
  
  pos <- rep(NaN,nrow(df))
  
  pos[st=="d3.s"&tr=="ACTnoOVA"] <- 1
  pos[tr=="OVAnoACT"] <- 2
  pos[st=="d3.s"&tr=="control"] <- 3
  pos[st=="d3.s"&tr=="mAB"] <- 4
  pos[st=="d7.s"&tr=="control"] <- 5
  pos[st=="d7.s"&tr=="mAB"] <- 6
  
  df$plot.pos <- pos
  df
  
  
}

#c(expression(ACT~(d3,OVA^-)),expression)

vdf <- relabl(vdf)
mdf <- relabl(mdf)

mdf$treated<-FALSE
mdf$treated[mdf$start=="d3.s"&mdf$tmid>3] <- TRUE
mdf$treated[mdf$start=="d7.s"&mdf$tmid>7] <- TRUE

vdf$treated<-FALSE
vdf$treated[vdf$start=="d3.s"&vdf$day>3] <- TRUE
vdf$treated[vdf$start=="d7.s"&vdf$day>7] <- TRUE


vdf$treated2<-FALSE
vdf$treated2[vdf$start=="d3.s"&vdf$day>=3] <- TRUE
vdf$treated2[vdf$start=="d7.s"&vdf$day>=6] <- TRUE

to_string <- as_labeller(c('1' = "ACT~OVA^'-'~(d3)", 
                           '2' = "no~ACT~OVA^'+'~(d3)",
                           '3' = "ACT-only~(d3)", 
                           '4' = "ACT+mAb~(d3)",
                           '5' = "ACT-only~(d7)", 
                           '6' = "ACT+mAb~(d7)"),label_parsed)

pg <-ggplot(mdf,aes(x=tmid,y=g,color=treated))+
  facet_wrap(~plot.pos,ncol=2,dir="h",labeller = to_string)+
  geom_abline(slope=0,intercept=0.5,color="grey70")+
  geom_point(aes(group=mouse))+
  geom_line(aes(group=mouse))+
  scale_x_continuous("days", breaks=seq(0,15,3))+
  scale_y_continuous(expression(paste("tumour volume growth", (day^-1))))+
  scale_color_manual("",labels=c("no ACT","ACT"), values=c("#000000","#009E73"))+
  theme_classic(base_size = 20)+
  theme(text = element_text(size=9))
pg

pv <-ggplot(vdf,aes(x=day))+
  facet_wrap(~plot.pos,ncol=2,dir="h",labeller = to_string)+
  geom_line(aes(y=0.01*exp(day/2)),color="grey70")+
  geom_line(aes(y=0.0025*exp(day/2)),color="grey70")+
  geom_line(aes(y=0.04*exp(day/2)),color="grey70")+
  geom_line(aes(group=mouse,y=vol,color=treated2))+
  geom_point(aes(group=mouse,y=vol,color=treated))+
  scale_x_continuous("days", breaks=seq(0,15,3))+
  scale_y_log10(expression(paste("tumour volume", (mm^3))))+
  scale_color_manual("",labels=c("no ACT","ACT"), values=c("#000000","#009E73"))+
  theme_classic(base_size = 20)+
  theme(text = element_text(size=9))+
  guides(color=FALSE)
pv



make_growth_lines <- function(g){
  day <- seq(0,14,1)
  data.frame(day=rep(day,4),vol=c(0.01*exp(day*g),0.0025*exp(day*g),0.04*exp(day*g),0.16*exp(day*g)),g=g,id=c(rep(1,length(day)),rep(2,length(day)),rep(3,length(day)),rep(4,length(day))))  
}

growth.lines <- do.call(rbind,lapply(c(0.3,0.4,0.5,0.6,0.7,0.8),make_growth_lines))

psx <-ggplot(growth.lines,aes(x=day,y=vol))+
  facet_grid(rows=vars(g))+
  geom_line(aes(group=id),color="grey70")+
  geom_line(data=subset(vdf,treatment=="OVAnoACT"),aes(group=mouse,y=vol))+
  geom_point(data=subset(vdf,treatment=="OVAnoACT"),aes(group=mouse,y=vol))+
 scale_x_continuous("days", breaks=seq(0,15,3))+
  scale_y_log10(expression(paste("tumour volume", (mm^3))))+
  theme_classic(base_size = 20)+
  theme(text = element_text(size=10),legend.position = "top")
psx

ggsave("/home/richard/Documents/03_CTLsGoesNative/000_archive/figures/s2.tiff",plot = psx,width = unit(3,"in"),height=unit(6.5,"in"),dpi=300)

```
REVISED CODE FOR KILLING PARAMETER TESTING

```{r,echo=FALSE}
## log likelihood function for Poisson distributions - intensities are either proportional to the number of CTLs, or TCs
ll.killing <- function(k,data,f.method="linear"){

  data$killing <- data$cTC_A + data$uTC_A
  if(f.method=="linear"){
    ## expected number of kills in each window 
    lambdas <- k*data$CTL*data$min/(24*60) 
    ll <- sum(sapply(1:length(lambdas), function(i) dpois(data$killing[i],lambdas[i],log=TRUE)))
} 
  
  if(f.method=="mass-action"){
    ## expected number of kills in each window 
    lambdas <- k*data$TC*data$CTL*data$min/(24*60) 
    ll <- sum(sapply(1:length(lambdas), function(i) dpois(data$killing[i],lambdas[i],log=TRUE)))
  } 
  return(ll)
}

#datadir <- "/home/richard/Documents/03_CTLsGoesNative/01_draft_0/01_processed_data"
datadir <- paste(base_dir,"data",sep="/")
setwd(datadir)
rawdf <- read.csv("formattedIntravital.csv")

data.all <- rawdf
data.mab <- rawdf[rawdf$ID=="mAB",]
data.contr <- rawdf[rawdf$ID=="control",]

df1 <- data.frame(model="linear",id="all",pr=exp(sapply(seq(0.2,0.8,0.005),ll.killing,data=data.all,f.method="linear")),k=seq(0.2,0.8,0.005))
df2 <- data.frame(model="linear",id="ACT+mAb",pr=exp(sapply(seq(0.2,0.8,0.005),ll.killing,data=data.mab,f.method="linear")),k=seq(0.2,0.8,0.005))
df3 <- data.frame(model="linear",id="ACT-only",pr=exp(sapply(seq(0.2,0.8,0.005),ll.killing,data=data.contr,f.method="linear")),k=seq(0.2,0.8,0.005))

df4 <- data.frame(model="mass-action",id="all",pr=exp(sapply(seq(0.1,0.5,0.005)/1000,ll.killing,data=data.all,f.method="mass-action")),k=seq(0.1,0.5,0.005)/1000)
df5 <- data.frame(model="mass-action",id="ACT+mAb",pr=exp(sapply(seq(0.1,0.5,0.005)/1000,ll.killing,data=data.mab,f.method="mass-action")),k=seq(0.1,0.5,0.005)/1000)
df6 <- data.frame(model="mass-action",id="ACT-only",pr=exp(sapply(seq(0.1,0.5,0.005)/1000,ll.killing,data=data.contr,f.method="mass-action")),k=seq(0.1,0.5,0.005)/1000)

df <- rbind(df1,df2,df3,df4,df5,df6)

p <-ggplot(df,aes(x=k,y=pr))+
  facet_grid(rows=vars(id),cols=vars(model),scales="free")+
  geom_line()
p


```

It is tricky to compare the likelihood scores between the the contacted-only and all-kills observations, because there are different numbers of available events. 
Case A) comparing contacted-only vs all-kills, there are simply more observed killing events in all-kills. Therefore, the probability of observing EXACTLY n kills is lower. 
Case B) if we try to include apoptosis in the contacted-only group, again it is not fair. Now we have to calculate the JOINT probability of EXACTLY n1 kills AND n2 apoptosis events.

Conclusion: we should be careful what we use these data to conclude!
Conclusion! Based on these data, the best choice for a killing model could be EITHER:

linear killing - no difference between days or treatments
dual-saturation - Either no difference between days or treatments, or grouped by day

Either way, none of these models favoured discriminating between control or mAB.

Best thing to do with this data: leave out the dual-sat fit, compare only mass-action and constant.

THIS MIGHT BE WRONG (CODE IN FOLLOWING CHUNK), BECAUSE IF WE SPLIT THE DATA UP THEN WE NEED TO ADD AN EXTRA (HYPER)PARAMETER FOR EACH SPLIT, I RECKON.
WE ARE GOING TO AVOID WORRYING ABOUT ANY OF THIS BY ONLY COMPARING THE MASS ACTION AND LINEAR FIT, AND ONLY LOOKING AT THE CASE WHERE WE FIT SEPARATE KILLING PARAMETERS PER GROUP. THAT IS BECAUSE WE WERE ONLY ASKED TO SHOW THAT THE NUMBER OF TUMOUR CELLS DOES NOT INFLUENCE THE KILLING RATE. 
```{r,echo=FALSE}
require(ggh4x)
require(reshape2)
require(ggplot2)
ll.apopto <- function(k,data){
      ## expected number of tumour cells apoptosis
      lambdas <- k*data$TC*data$min/(24*60) 
      ll <- sum(sapply(1:length(lambdas), function(i) dpois(data$uTC_A[i],lambdas[i],log=TRUE)))
      return(ll)
}

ll.killing <- function(par,data,f.method="linear"){
  par <- abs(par)
  k <- par[1]
  if(f.method=="linear"){
    ## expected number of kills in each window 
    lambdas <- k*data$CTL*data$min/(24*60) 
    ll <- sum(sapply(1:length(lambdas), function(i) dpois(data$killing[i],lambdas[i],log=TRUE)))
} 
  
  if(f.method=="mass-action"){
    ## expected number of kills in each window 
    lambdas <- k*data$TC*data$CTL*data$min/(24*60) 
    ll <- sum(sapply(1:length(lambdas), function(i) dpois(data$killing[i],lambdas[i],log=TRUE)))
  } 
  
  if(f.method=="dual-sat"){
    he <- par[2]
    ht <- par[3]
    lambdas <- k*data$TC*data$CTL*data$min/(24*60) * 1/(1+he*data$CTL+ht*data$TC)
    ll <- sum(sapply(1:length(lambdas), function(i) dpois(data$killing[i],lambdas[i],log=TRUE)))
  }
  
  return(ll)
  
}

fit.killing <- function(subset,data,method="linear",apopto=FALSE){
  ## subset is a vector with first element representing the treatment, second element representing the day
  if(subset[1]!="all"){
    data <- data[data$ID==subset[1],]
  }
  if(subset[2]!="all"){
    data <- data[data$day==subset[2],]
  }
  
  if(subset[3]=="contacted") data$killing <- data$cTC_A
  if(subset[3]=="all") data$killing <- data$cTC_A + data$uTC_A
  
  
  if(method=="linear" | method=="mass-action"){
    fit <- optimize(ll.killing,interval=c(0,10),data=data,f.method=method,maximum=TRUE)
    fit <- data.frame(treatment=subset[1],day=subset[2],contacted=subset[3],k=fit$maximum,log.lik=fit$objective)
  }
  
  if(method=="dual-sat"){
    fit <- optim(c(1,1,1),ll.killing,data=data,f.method="dual-sat",control=list(fnscale=-1))
    fit <- data.frame(treatment=subset[1],day=subset[2],contacted=subset[3],k=fit$par[1],he=fit$par[2],ht=fit$par[3],log.lik=fit$value)
  }
  
  if(apopto & subset[3] == "contacted"){
    f.apt <- optimize(ll.apopto,interval=c(0,1),data=data,maximum=TRUE)
    fit$ka=f.apt$maximum
    fit$log.lik=fit$log.lik+f.apt$objective
  }
  
  return(fit)
  
}

make.subsets <- function(contact){
  subsets <- list(c("all","all",contact),
                c("mAB","all",contact),
                c("control","all",contact),
                c("all","6",contact),
                c("mAB","6",contact),
                c("control","6",contact),
                c("all","9",contact),
                c("mAB","9",contact),
                c("control","9",contact))
  return(subsets)
}

fit.models <- function(rawdf,contact,return.k=FALSE){
  
  subsets <- make.subsets(contact)
  groupings <- list(none=c(1),
                  by.day=c(4,7),
                  by.treatment=c(2,3),
                  by.both=c(5,6,8,9))
  
  fit.linear <- do.call(rbind,lapply(subsets,fit.killing,data=rawdf,method="linear"))
  linear <- sapply(groupings, function(g) sum(fit.linear[g,'log.lik']))
  k.linear <- sapply(groupings, function(g) sum(fit.linear[g,'log.lik']))

  fit.mass <- do.call(rbind,lapply(subsets,fit.killing,data=rawdf,method="mass-action"))
  mass <- sapply(groupings, function(g) sum(fit.mass[g,'log.lik']))

  fit.ds <- do.call(rbind,lapply(subsets,fit.killing,data=rawdf,method="dual-sat"))
  ds <- sapply(groupings, function(g) sum(fit.ds[g,'log.lik']))
  
  if(return.k) return(list(linear=fit.linear,mass=fit.mass))
  
  npar <- c(1,1,3) %*% t(sapply(groupings,length))
  perf <- rbind(linear,mass,ds)
  aic <- 2*(npar-perf)
  n <- nrow(rawdf)
  bic <- log(n)*npar-2*perf
  
  perf <- melt(perf)
  names(perf) <- c('model','subset','log.lik')
  perf$aic <- melt(aic)$value
  perf$bic <- melt(bic)$value
  perf$contact <- rep(contact,nrow(perf))
  return(perf)
}

#datadir <- "/home/richard/Documents/03_CTLsGoesNative/01_draft_0/01_processed_data"
datadir <- paste(base_dir,"data",sep="/")
setwd(datadir)
rawdf <- read.csv("formattedIntravital.csv")
rawdf$day <- sapply(rawdf$day,toString)

rawdf <- rawdf[,c('ID','min','day','CTL','TC','cTC_A','uTC_A')]
rawdf[is.na(rawdf)]<-0


all.kills <- fit.models(rawdf,"all")
all.kills <- subset(all.kills,model!="ds")
contact.kills <- fit.models(rawdf,"contacted")
contact.kills <- subset(contact.kills,model!="ds")

all.kills <- melt(all.kills,measure.vars=c("aic","bic"))
df <- subset(all.kills,variable=="bic")
p <- ggplot(df,aes(x=model,y=value,shape=subset,color=subset))+
  geom_point(size=3)+
  scale_color_viridis_d()+
  scale_y_continuous("bic")+
  theme_classic(base_size=20)
p

df <- subset(all.kills,variable=="aic")
p <- ggplot(df,aes(x=model,y=value,shape=subset,color=subset))+
  geom_point(size=3)+
  scale_color_viridis_d()+
    scale_y_continuous("aic")+
  theme_classic(base_size=20)
p

contact.kills <- melt(contact.kills,measure.vars=c("aic","bic"))
df <- subset(contact.kills,variable=="bic")
p <- ggplot(df,aes(x=model,y=value,shape=subset,color=subset))+
  geom_point(size=3)+
  scale_color_viridis_d()+
  scale_y_continuous("bic")+
  theme_classic(base_size=20)
p

df <- subset(contact.kills,variable=="aic")
p <- ggplot(df,aes(x=model,y=value,shape=subset,color=subset))+
  geom_point(size=3)+
  scale_color_viridis_d()+
  scale_y_continuous("aic")+
  theme_classic(base_size=20)
p

ss <- all.kills[all.kills$subset%in%c("none","by.treatment"),]

labz <- as_labeller(c('none'="all samples",'by.treatment'="by treatment",'aic'="Akaike",'bic'="Bayes"))

p <- ggplot(ss,aes(x=model,y=value))+
  facet_grid(rows=vars(variable),cols=vars(subset),labeller = labz,scales="free")+
  geom_point(size=3)+
  scale_color_viridis_d()+
  scale_y_continuous("information criterion value")+
  scale_x_discrete("",labels=c("linear","mass-action"))+
  theme_classic(base_size=20)+ 
  theme(panel.spacing = unit(3, "lines"))
p


ps2a <- ggplot(ss,aes(x=value,y=model))+
  facet_nested(variable+subset~.,labeller = labz,scales="free")+
  #facet_grid(rows=vars(variable),cols=vars(subset),labeller = labz,scales="free")+
  geom_point(size=3)+
  scale_color_viridis_d()+
  scale_x_continuous("information criterion value")+
  scale_y_discrete("",labels=c("linear","mass-action"))+
  theme_classic(base_size=12)+ 
  theme(panel.spacing = unit(3, "lines"))

ps2a

```

Retrieve best fitting killing rates for linear and mass-action killing

```{r,echo=FALSE}
all.kills <- fit.models(rawdf,"all",return.k = TRUE)

linear <- all.kills$linear
linear <- linear[linear$day=="all",c("treatment","k")]
linear$type <- "linear"

mass <- all.kills$mass
mass <- mass[mass$day=="all",c("treatment","k")]
mass$type <- "mass-action"

df <- rbind(mass,linear)

df
            

```
Assuming the values of k are the intensity parameters for a poisson distribution, calculate the expected value and 1-99% confidence intervals:

```{r, echo=FALSE}
build.fit <- function(i,df){
  trt <- df$treatment[i]
  type <- df$type[i]
  k <- df$k[i]/24
  ctls <- seq(0,1000,1)
  if(type=="linear") exp_ctl_pois <- data.frame(X=ctls,conf05=sapply(ctls, function(x) qpois(0.01,lambda=k*x)),
                           conf95=sapply(ctls, function(x) qpois(0.99,lambda=k*x)),
                           exp=sapply(ctls, function(x) qpois(0.5,lambda=k*x)),k=k,trt=trt,type=type)
  if(type=="mass-action"){
    ctls <- ctls*1400
    exp_ctl_pois <- data.frame(X=ctls,conf05=sapply(ctls, function(x) qpois(0.01,lambda=k*x)),
                           conf95=sapply(ctls, function(x) qpois(0.99,lambda=k*x)),
                           exp=sapply(ctls, function(x) qpois(0.5,lambda=k*x)),k=k,trt=trt,type=type)
  }
  return(exp_ctl_pois)
}


fits <- lapply(1:nrow(df),build.fit,df=df)
fits <- do.call(rbind,fits)
```


retrieve the killing statistics from the data:
```{r, echo=FALSE}

rawdf <- read.csv(paste(base_dir,"/data/formattedIntravital.csv",sep=""))

lin <- data.frame(type="linear",trt=rawdf$ID,X=rawdf$CTL,kills=60*(rawdf$cTC_A+rawdf$uTC_A)/(rawdf$min))
lin2 <- lin
lin2$trt <- "all"
lin <- rbind(lin,lin2)

massa <- data.frame(type="mass-action",trt=rawdf$ID,X=rawdf$CTL*rawdf$TC,kills=60*(rawdf$cTC_A+rawdf$uTC_A)/(rawdf$min))
massa2 <- massa
massa2$trt <- "all"
massa <- rbind(massa,massa2)

```


```{r}

labz <- as_labeller(c('all' = "all samples", 
                           'mAB' = "ACT+mAb",
                           'control' = "ACT-only"))



ps2b <- ggplot(subset(fits,type=="linear"),aes(x=X))+
  facet_grid(cols=vars(trt),scales = "free",labeller=labz)+
  geom_point(data=lin,aes(y=kills))+
  geom_ribbon(aes(ymin=conf05,ymax=conf95),alpha=0.1)+
  geom_line(aes(y=exp),color="red")+
  #geom_point(size=1.5,aes(color=treatment,y=TCa*TC/24))+
  theme_classic(base_size=12)+
  scale_x_continuous(expression(CTLs~at~position))+
  scale_y_continuous(expression(TC~apoptosis~events~(hour^-1)))
ps2b

ps2c <- ggplot(subset(fits,type=="mass-action"),aes(x=X))+
  facet_grid(cols=vars(trt),scales = "free",labeller=labz)+
    geom_point(data=massa,aes(y=kills))+
  geom_ribbon(aes(ymin=conf05,ymax=conf95),alpha=0.1)+
  geom_line(aes(y=exp),color="red")+
  #geom_point(size=1.5,aes(color=treatment,y=TCa*TC/24))+
  theme_classic(base_size=12)+
  scale_x_continuous(expression(product~of~tumour~cells~and~CTLs~at~position))+
  scale_y_continuous(expression(TC~apoptosis~events~(hour^-1)))
ps2c



#ps1 <- egg:
```

Arrange plots for figure S1

```{R}
require(cowplot)
fs2 <- ggdraw()+
  draw_plot(ps2a,0,0,width=0.4,height=1)+
  draw_plot(ps2b,0.4,0.5,width=0.6,height=0.5)+
  draw_plot(ps2c,0.4,0,width=0.6,height=0.5)+
  draw_plot_label(c("A","B","C"),x=c(0,0.4,0.4),y=c(1,1,0.5))

ggsave("/home/richard/Documents/03_CTLsGoesNative/000_archive/figures/s1.tiff",plot = fs2,width = unit(7.5,"in"),height=unit(6,"in"),dpi=300)

```

```{r, echo=FALSE}

rawdf <- read.csv(paste(base_dir,"/data/formattedIntravital.csv",sep=""))
lambda <- 60*sum(rawdf$cTC_A+rawdf$uTC_A)/sum(rawdf$CTL*rawdf$min) ## estimated hourly killing rate per CTL
ctls <- seq(0,1000,1)
exp_ctl_pois <- data.frame(CTL=ctls,conf05=sapply(ctls, function(x) qpois(0.01,lambda=lambda*x)),
                           conf95=sapply(ctls, function(x) qpois(0.99,lambda=lambda*x)),
                           exp=sapply(ctls, function(x) qpois(0.5,lambda=lambda*x)))

require(ggplot2)
data_proc_funcs <- paste(base_dir,"/rscripts/utility/data_processing_functions.R",sep="")
source (data_proc_funcs)
iv.proc <- get.iv()
iv.proc$treatment <- "ACT-only"
iv.proc$treatment[iv.proc$ID=="mAB"]<- "ACT+mAb"

measure_mit_funcs <- paste(base_dir,"/rscripts/utility/measure_mitosis.R",sep="")

source(measure_mit_funcs)
print(summary)

df.smm <- rbind(data.frame(CTL=iv.proc$CTL,TC=iv.proc$TC,mitosis=iv.proc$TC*iv.proc$TCm/24,apoptosis=iv.proc$TC*iv.proc$TCa/24,day=iv.proc$day,treatment=iv.proc$treatment,start="d3.s"),
                data.frame(CTL=summary$CTL,TC=summary$TCs,mitosis=summary$mitosis,apoptosis=NA,day=10,treatment=summary$treatment,start="d7.s"))




p2c <- ggplot(iv.proc,aes(x=CTL))+
  geom_ribbon(data=exp_ctl_pois,aes(ymin=conf05,ymax=conf95),alpha=0.1)+
  geom_line(data=exp_ctl_pois,aes(y=exp,linetype="Poisson"),color="red")+
  geom_point(size=1.5,aes(color=treatment,y=TCa*TC/24))+
  theme_classic(base_size=20)+
  scale_x_continuous("total CTLs at position")+
  scale_y_continuous(expression(TC~apoptosis~events~(hour^-1)))+
  scale_linetype("",labels=c("Poisson \nexpectation"))+
  theme(text = element_text(size=10),legend.position = "top",legend.direction = "vertical")
p2c

p2d <- ggplot(df.smm,aes(x=CTL,y=mitosis))+
  geom_point(size=1.5,aes(color=treatment,shape=start))+
  theme_classic(base_size=20)+
  scale_x_continuous("total CTLs at position")+
  scale_y_continuous(expression(TC~mitosis~events~(hour^-1)))+
  scale_shape("treatment \nstart",labels=c("day 3", "day 7"))+
  theme(text = element_text(size=10),legend.position = "top",legend.direction = "vertical")
p2d

dfx <- reshape2::melt(df.smm,measure.vars=c("apoptosis","mitosis"))
exp_ctl_pois$variable<- "apoptosis"
p2cd <- ggplot(dfx,aes(x=CTL))+
  facet_grid(cols=vars(variable))+
  geom_ribbon(data=exp_ctl_pois,aes(ymin=conf05,ymax=conf95),alpha=0.1)+
  geom_line(data=exp_ctl_pois,aes(y=exp,linetype="Poisson"),color="red")+
  geom_point(size=1.5,aes(y=value,color=treatment,shape=start))+
  theme_classic(base_size=20)+
  scale_x_continuous("total CTLs at position")+
  scale_y_continuous(expression(events~(TC^-1~hour^-1)))+
  scale_shape("treatment \nstart",labels=c("day 3", "day 7"))+
    scale_linetype("",labels=c("Poisson \nexpectation"))+
  theme_classic(base_size=20)+
  theme(text = element_text(size=10))
p2cd



```

```{r}
f2_top <- egg::ggarrange(pv,pg,nrow=1,labels=c("A","B"))
f2_bot <- egg::ggarrange(p2cd,nrow=1,labels=c("C"))

#f2 <- egg::ggarrange(pv,pg,p2c,p2d,nrow=2,labels=c("A","B","C","D"),heights=c(3,1))

ggsave("/home/richard/Documents/03_CTLsGoesNative/000_archive/figures/f2ab.tiff",plot = f2_top,width = unit(7.5,"in"),height=unit(6,"in"),dpi=300)
ggsave("/home/richard/Documents/03_CTLsGoesNative/000_archive/figures/f2c.tiff",plot = f2_bot,width = unit(7.5,"in"),height=unit(3.5,"in"),dpi=300)

```

```{r}
require(ggplot2)
filesplitter <- function(filename){
  x <- strsplit(filename,split=".csv")[[1]]
  x <- strsplit(x,split="_")[[1]]
  data.frame(mouse=x[1],day=x[2],position=x[3])
}

process_imdata <- function(foldername){
  ctl_data <- lapply(list.files(foldername), function(ff) read.csv(paste(foldername,ff,sep="/")))
  ctl_ids <- lapply(list.files(foldername), filesplitter)
  
  no_observations <- which(sapply(ctl_data,nrow)==0)
  
  if(length(no_observations)>0){
    ctl_data <- ctl_data[-no_observations]
    ctl_ids <- ctl_ids[-no_observations]
  }
  
  ctl_data <- lapply(1:length(ctl_data), function(i) cbind(ctl_data[[i]],ctl_ids[[i]]))
  do.call(rbind,ctl_data)
}

datadir <- paste(base_dir,"data/manual_image_counting",sep="/")
setwd(datadir)

ctl_data <- process_imdata("d7s_CTLs")
ctl_data$type <- "ctl"
mitosis_data <- process_imdata("d7s_mitosis")
mitosis_data$type <- "mitosis"

dat <- rbind(ctl_data,mitosis_data)

lablr <- function(x){
  c("ACT+mAb (m48)","ACT-only (m51)","ACT-only (m52)","ACT-only (m53)","ACT+mAb (m54)","ACT+mAb (m56)",
    "p00","p01","p02","p03","p04")[x==c("m48","m51","m52","m53","m54","m56","p00","p01","p02","p03","p04")]
}

lablr <- as_labeller(c('m48' = "ACT+mAb~(m48)", 'm51'="ACT-only~(m51)", 'm52'="ACT-only~(m52)", 'm53'="ACT-only~(m53)",  'm54' = "ACT+mAb~(m54)",'m56' = "ACT+mAb~(m56)",
                       'p00'="p00",'p01'="p01",'p02'="p02",'p03'="p03",'p04'="p04"),label_parsed)

dat$treatment <- "ACT-only"
dat$treatment[dat$mouse%in%c("m48","m54","m56")] <- "ACT+mAb"

get_legend<-function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
}

pa <- ggplot(dat,aes(x=X*0.7,y=Y*0.7,color=type))+
  facet_grid(position~treatment+mouse)+
  geom_point(alpha=0.6)+
  scale_color_viridis_d(labels=c("CTL","mitosis"))+
  theme_classic(base_size = 12)+
  scale_x_continuous(expression(X~(paste(mu,m,sep=""))))+
  scale_y_continuous(expression(Y~(paste(mu,m,sep=""))))
s6_leg <- get_legend(pa)
pa <- pa+theme(legend.position = "none")

pa

pb <- ggplot(dat,aes(x=X*0.7,y=Slice*7,color=type))+
  facet_grid(position~treatment+mouse)+
  geom_point(alpha=0.6)+
  scale_color_viridis_d(labels=c("CTL","mitosis"))+
  theme_classic(base_size = 12)+
  scale_x_continuous(expression(X~(paste(mu,m,sep=""))))+
  scale_y_continuous(expression(Z~(paste(mu,m,sep=""))),breaks=c(0,50,100))+
  theme(legend.position = "none")
pb

pc <- ggplot(dat,aes(x=Slice*7,y=Y*0.7,color=type))+
  facet_grid(position~treatment+mouse)+
  geom_point(alpha=0.6)+
  scale_color_viridis_d(labels=c("CTL","mitosis"))+
  theme_classic(base_size = 12)+
  scale_x_continuous(expression(Z~(paste(mu,m,sep=""))),breaks=c(0,50,100))+
  scale_y_continuous(expression(Y~(paste(mu,m,sep=""))))+
  theme(legend.position = "none")
pc

require(cowplot)
ps6 <- ggdraw() +
 draw_plot(pa, 0, 0.3, 0.7, 0.7) +
 draw_plot(pc, 0.7, 0.3, 0.3, .7) +
  draw_plot(pb, 0,0,0.7,0.3)+
  draw_plot(s6_leg,0.7,0.1,0.2,0.2)+
  draw_plot_label(c("A", "B", "C"), c(0, 0.7, 0), c(1, 1, 0.3), size = 15)
ggsave("/home/richard/Documents/03_CTLsGoesNative/000_archive/figures/fs6.tiff",plot = ps6,width = unit(18,"in"),height=unit(12,"in"),dpi=300)
```


